---
title: 'Task 1: Oil spills in CA (PELEGRI)'
author: "Patrick Pelegri-O'Day"
date: "3/1/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(raster)
library(tidyverse)
library(here)
library(broom)
library(sf)
library(tmap)
library(spatstat)
library(maptools)
```

### Overview

This report visualizes data on oil spills in California recorded in 2008 in the Oil Spill Prevention and Response (OSPR) Incident Tracking database. An oil spill incident in this dataset is defined as "a discharge or threatened discharge of petroleum or other deleterious material into the waters of the state."

**Citation:** Oil Spill Incident Tracking [ds394], 2009-07-23. 2008 Edition.

### Setup

Read in data
```{r}
oil_raw_sf <- read_sf(here('data', 'ds394', 'ds394.shp'))

oil_sf <- oil_raw_sf %>% 
  janitor::clean_names()

counties_sf <- read_sf(here('data/counties/CA_Counties_TIGER2016.shp'))

counties_subset_sf <- counties_sf %>% 
  janitor::clean_names() %>% 
  select(county_name = name, land_area = aland)
```

Set coordinate system
```{r}
# st_crs(oil_sf) # EPSG is 4269
# 
# st_crs(counties_subset_sf) # EPSG is 4326

### Projections are different. To align them, we take the CA county projection and apply it to the oil spill data.

oil_4326_sf <- st_transform(oil_sf, st_crs(counties_subset_sf))
```

### Visual analysis

Plot with tmap
```{r}
tmap_mode(mode = 'view')
tm_shape(counties_subset_sf) + 
  tm_borders() +
  tm_fill('land_area', 
          title = "Land Area (meters squared)",
          palette = 'BuGn') +
  tm_borders(col = 'black') +
tm_shape(oil_4326_sf) +
  tm_dots(col = 'darkslateblue')
```

**Figure 1:** This interactive map shows oil spills recorded in California in 2008 by the California Department of Fish and Wildlife.

Make chloropleth map based on count of inland spill events by county
```{r}
# Join the spill sf and county sf
county_spills_sf <- counties_subset_sf %>% 
  st_join(oil_4326_sf)

# Only retain inland spills
inland_spills_sf <- county_spills_sf %>% 
  filter(inlandmari == 'Inland') %>% 
  group_by(county_name) %>% 
  summarize(spill_count = sum(!is.na(dfgcontrol)))

ggplot(data = inland_spills_sf) +
  geom_sf(aes(fill = spill_count), color = 'white', size = 0.1) + 
  scale_fill_gradientn(colors = c('lightgrey', 'darkgoldenrod2', 'darkgoldenrod4')) +
  theme_minimal() + 
  labs(fill = 'Number of inland oil spills')
```

**Figure 2:** Map of inland oil spills by county in 2008. Counties with a greater number of spills are shaded with a darker color.

Combine the two datasets as a spatial point pattern
```{r}
# Convert oil spills sf into a spatial point pattern
oil_sp <- as(oil_4326_sf, 'Spatial')
oil_ppp <- as(oil_sp, 'ppp')

# Convert counties sf into a window
counties_sp <- as(counties_subset_sf, 'Spatial')
counties_win <- as(counties_sp, 'owin')

# Combine
oil_counties_ppp <- ppp(oil_ppp$x, oil_ppp$y, window = counties_win)
```

Make the G function
```{r}
# QUESTION: What is the significance of having difference distances?
r_vec <- seq(0, 10000, by = 100) 
 
# Define G function
gfunction <- envelope(oil_counties_ppp, fun = Gest, r = r_vec, nsim = 10, nrank = 2) 
```

```{r}
# Reformat for plotting
gfunction_long <- gfunction %>% 
  as.data.frame() %>% 
  pivot_longer(cols = obs:hi, names_to = "model", values_to = "g_val") %>% 
  mutate(model = case_when(
          model == "hi" ~ "High",
          model == 'lo' ~ 'Low',
          model == 'obs' ~ 'Observed',
          model == 'theo' ~ 'Theoretical'))
```

```{r}
# Plot G function
ggplot(data = gfunction_long, aes(x = r, y = g_val, group = model)) +
  geom_line(aes(color = model)) +
  theme_minimal() +
  labs(x = "Distance", y = "G value", color = "Model")
```

**Figure 3:** This plot shows that oil spills are significantly more clustered than they would be under complete spatial randomness based on 10 iterations of a G function.